# Clustering with k-means

## k-means

## Example: Finding Teen Market Segments

### Step 2: Exploring and preparing the data

```{r}
teens <- read.csv("Chapter 09/snsdata.csv")
str(teens)
```

- look at missing data for female variable

```{r}
table(teens$gender)
table(teens$gender, useNA = "ifany")
```

- look at missing data for age variable

```{r}
summary(teens$age)
```

- eliminate age outliers

```{r}
teens$age <- ifelse(teens$age >= 13 & teens$age < 20,
                     teens$age, NA)

summary(teens$age)
```

- reassign missing gender values to "unknown"

```{r}
teens$female <- ifelse(teens$gender == "F" &
                         !is.na(teens$gender), 1, 0)
teens$no_gender <- ifelse(is.na(teens$gender), 1, 0)
```

- check our recoding work

```{r}
table(teens$gender, useNA = "ifany")
table(teens$female, useNA = "ifany")
table(teens$no_gender, useNA = "ifany")
```

- finding the mean age by cohort

```{r}
mean(teens$age) # doesn't work
mean(teens$age, na.rm = TRUE) # works
```

- age by cohort

```{r}
aggregate(data = teens, age ~ gradyear, mean, na.rm = TRUE)
```

- create a vector with the average age for each gradyear, repeated by person

```{r}
ave_age <- ave(teens$age, teens$gradyear,
                 FUN = function(x) mean(x, na.rm = TRUE))


teens$age <- ifelse(is.na(teens$age), ave_age, teens$age)
```

- check the summary results to ensure missing values are eliminated

```{r}
summary(teens$age)
```

### Step 3: Training a model on the data

```{r}
interests <- teens[5:40]
interests_z <- as.data.frame(lapply(interests, scale))

set.seed(2345)
teen_clusters <- kmeans(interests_z, 5)
```

### Step 4: Evaluating model performance

- look at the size of the clusters

```{r}
teen_clusters$size
```

- look at the cluster centers

```{r}
teen_clusters$centers
```

### Step 5: Improving model performance

- apply the cluster IDs to the original data frame

```{r}
teens$cluster <- teen_clusters$cluster
```

- look at the first five records

```{r}
teens[1:5, c("cluster", "gender", "age", "friends")]
```

- mean age by cluster

```{r}
aggregate(data = teens, age ~ cluster, mean)
```

- proportion of females by cluster

```{r}
aggregate(data = teens, female ~ cluster, mean)
```

- mean number of friends by cluster

```{r}
aggregate(data = teens, friends ~ cluster, mean)
```
