# Managing and Understanding Data

```{r setup, include=FALSE}
library(gmodels)
```

## R data structures

### Vectors

- create vectors of data for three medical patients

```{r}
subject_name <- c("John Doe", "Jane Doe", "Steve Graves")
temperature <- c(98.1, 98.6, 101.4)
flu_status <- c(FALSE, FALSE, TRUE)
```

- access the second element in body temperature vector

```{r}
temperature[2]
```

- examples of accessing items in vector
- include items in the range 2 to 3

```{r}
temperature[2:3]
```

- exclude item 2 using the minus sign

```{r}
temperature[-2]
```

- use a vector to indicate whether to include item

```{r}
temperature[c(TRUE, TRUE, FALSE)]
```

### Factors

- add gender factor

```{r}
gender <- factor(c("MALE", "FEMALE", "MALE"))
gender
```

- add blood type factor

```{r}
blood <- factor(c("O", "AB", "A"),
                levels = c("A", "B", "AB", "O"))
blood
```

- add ordered factor

```{r}
symptoms <- factor(c("SEVERE", "MILD", "MODERATE"),
                   levels = c("MILD", "MODERATE", "SEVERE"),
                   ordered = TRUE)
symptoms
```

- check for symptoms greater than moderate

```{r}
symptoms > "MODERATE"
```

### Lists

- display information for a patient

```{r}
subject_name[1]
temperature[1]
flu_status[1]
gender[1]
blood[1]
symptoms[1]
```

- create list for a patient

```{r}
subject1 <- list(fullname = subject_name[1], 
                 temperature = temperature[1],
                 flu_status = flu_status[1],
                 gender = gender[1],
                 blood = blood[1],
                 symptoms = symptoms[1])
```

- display the patient

```{r}
subject1
```

- methods for accessing a list

- get a single list value by position (returns a sub-list)

```{r}
subject1[2]
```

- get a single list value by position (returns a numeric vector)

```{r}
subject1[[2]]
```

- get a single list value by name

```{r}
subject1$temperature
```

- get several list items by specifying a vector of names

```{r}
subject1[c("temperature", "flu_status")]
```

- access a list like a vector
- get values 2 and 3

```{r}
subject1[2:3]
```

### Data frames

- create a data frame from medical patient data

```{r}
pt_data <- data.frame(subject_name, temperature, flu_status, gender,
                      blood, symptoms, stringsAsFactors = FALSE)
```

- display the data frame

```{r}
pt_data
```

- accessing a data frame

- get a single column

```{r}
pt_data$subject_name
```

- get several columns by specifying a vector of names

```{r}
pt_data[c("temperature", "flu_status")]
```

- this is the same as above, extracting temperature and flu_status

```{r}
pt_data[2:3]
```

- accessing by row and column

```{r}
pt_data[1, 2]
```

- accessing several rows and several columns using vectors

```{r}
pt_data[c(1, 3), c(2, 4)]
```

- Leave a row or column blank to extract all rows or columns

- column 1, all rows

```{r}
pt_data[, 1]
```

- row 1, all columns

```{r}
pt_data[1, ]
```

- all rows and all columns

```{r}
pt_data[ , ]
```

- the following are equivalent

```{r}
pt_data[c(1, 3), c("temperature", "gender")]
pt_data[-2, c(-1, -3, -5, -6)]
```

### Matrixes

- create a 2x2 matrix

```{r}
m <- matrix(c(1, 2, 3, 4), nrow = 2)
m
```

- equivalent to the above

```{r}
m <- matrix(c(1, 2, 3, 4), ncol = 2)
m
```

- create a 2x3 matrix

```{r}
m <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2)
m
```

- create a 3x2 matrix

```{r}
m <- matrix(c(1, 2, 3, 4, 5, 6), ncol = 2)
m
```

- extract values from matrixes

```{r}
m[1, 1]
m[3, 2]
```

- extract rows

```{r}
m[1, ]
```

- extract columns

```{r}
m[, 1]
```

## Managing data with R

### saving, loading, and removing R data structures

- show all data structures in memory

```{r}
ls()
```

- remove the m and subject1 objects

```{r}
rm(m, subject1)
ls()

rm(list=ls())
```

## Exploring and understanding data

## data exploration example using used car data

```{r}
usedcars <- read.csv("Chapter 02/usedcars.csv", stringsAsFactors = FALSE)
```

### Exploring the structure of data

- get structure of used car data

```{r}
str(usedcars)
```

### Exploring numeric variables

- summarize numeric variables

```{r}
summary(usedcars$year)
summary(usedcars[c("price", "mileage")])
```

- calculate the mean income

```{r}
(36000 + 44000 + 56000) / 3
mean(c(36000, 44000, 56000))
```

- the median income

```{r}
median(c(36000, 44000, 56000))
```

- the min/max of used car prices

```{r}
range(usedcars$price)
```

- the difference of the range

```{r}
diff(range(usedcars$price))
```

- IQR for used car prices

```{r}
IQR(usedcars$price)
```

- use quantile to calculate five-number summary

```{r}
quantile(usedcars$price)
```

- the 99th percentile

```{r}
quantile(usedcars$price, probs = c(0.01, 0.99))
```

- quintiles

```{r}
quantile(usedcars$price, seq(from = 0, to = 1, by = 0.20))
```

- boxplot of used car prices and mileage

```{r}
boxplot(usedcars$price, main="Boxplot of Used Car Prices",
      ylab="Price ($)")

boxplot(usedcars$mileage, main="Boxplot of Used Car Mileage",
      ylab="Odometer (mi.)")
```

- histograms of used car prices and mileage

```{r}
hist(usedcars$price, main = "Histogram of Used Car Prices",
     xlab = "Price ($)")

hist(usedcars$mileage, main = "Histogram of Used Car Mileage",
     xlab = "Odometer (mi.)")
```

- variance and standard deviation of the used car data

```{r}
var(usedcars$price)
sd(usedcars$price)
var(usedcars$mileage)
sd(usedcars$mileage)
```

## Exploring categorical variables

- one-way tables for the used car data

```{r}
table(usedcars$year)
table(usedcars$model)
table(usedcars$color)
```

- compute table proportions

```{r}
model_table <- table(usedcars$model)
prop.table(model_table)
```

- round the data

```{r}
color_table <- table(usedcars$color)
color_pct <- prop.table(color_table) * 100
round(color_pct, digits = 1)
```

### Exploring relationships between variables

- scatterplot of price vs. mileage

```{r}
plot(x = usedcars$mileage, y = usedcars$price,
     main = "Scatterplot of Price vs. Mileage",
     xlab = "Used Car Odometer (mi.)",
     ylab = "Used Car Price ($)")
```

- new variable indicating conservative colors

```{r}
usedcars$conservative <-
  usedcars$color %in% c("Black", "Gray", "Silver", "White")
```

- checking our variable

```{r}
table(usedcars$conservative)
```

- Crosstab of conservative by model

```{r}
CrossTable(x = usedcars$model, y = usedcars$conservative)
```