# Association Rules

```{r setup, include=FALSE}
library(arules)
```

## Market Basket Analysis

## Example: Identifying Frequently-Purchased Groceries

### Step 2: Exploring and preparing the data

- load the grocery data into a sparse matrix

```{r}
groceries <- read.transactions("Chapter 08/groceries.csv", sep = ",")
summary(groceries)
```

- look at the first five transactions

```{r}
inspect(groceries[1:5])
```

- examine the frequency of items

```{r}
itemFrequency(groceries[, 1:3])
```

- plot the frequency of items

```{r}
itemFrequencyPlot(groceries, support = 0.1)
itemFrequencyPlot(groceries, topN = 20)
```

- a visualization of the sparse matrix for the first five transactions

```{r}
image(groceries[1:5])
```

- visualization of a random sample of 100 transactions

```{r}
image(sample(groceries, 100))
```

### Step 3: Training a model on the data

- default settings result in zero rules learned

```{r}
apriori(groceries)
```

- set better support and confidence levels to learn more rules

```{r}
groceryrules <- apriori(groceries, parameter = list(support =
                          0.006, confidence = 0.25, minlen = 2))
groceryrules
```

### Step 4: Evaluating model performance

- summary of grocery association rules

```{r}
summary(groceryrules)
```

- look at the first three rules

```{r}
inspect(groceryrules[1:3])
```

### Step 5: Improving model performance

- sorting grocery rules by lift

```{r}
inspect(sort(groceryrules, by = "lift")[1:5])
```

- finding subsets of rules containing any berry items

```{r}
berryrules <- subset(groceryrules, items %in% "berries")
inspect(berryrules)
```

- writing the rules to a CSV file

```{r}
write(groceryrules, file = "Chapter 08/groceryrules.csv",
      sep = ",", quote = TRUE, row.names = FALSE)
```

- converting the rule set to a data frame

```{r}
groceryrules_df <- as(groceryrules, "data.frame")
str(groceryrules_df)
```